---
title: "Homework Assignment: Sentiment Analysis of Emma"
format: html
editor: visual
author: Inaya Ahmed
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 3, fig.width = 4)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

**Title**: Exploring Joyful Language in Jane Austen’s *Emma* using Tidytext

### **Instructions**

Complete the following exercises using the appropriate packages in R. Ensure that your solutions are optimized and use functional programming principles where applicable.

1.  Load the necessary libraries.
2.  Answer each question in separate R code chunks.
3.  Provide detailed explanations for your approach.
4.  Submit the rendered HTML file.

#### **Objective**:

Use the **`tidytext`** package and three different **sentiment lexicons** (`nrc`, `afinn`, `bing`) to explore **positive/joyful words** in *Emma* by Jane Austen. You will tokenize the text, apply sentiment filters, visualize frequent sentiment words using `ggplot2`, and create a word cloud.

### **Tasks**:

1.  **Data Preparation**

    -   Load the `austen_books()` dataset from the **`janeaustenr`** package.
    -   Group by book and detect chapter boundaries using regex.
    -   Create `linenumber` and `chapter` columns.

```{r}
# Loading libraries:
library(tidyverse)
library(tidytext)
library(janeaustenr)
library(stringr)
library(wordcloud)
library(RColorBrewer)
```

```{r}
# Grouping by book and detecting chapter boundaries:
austen <- austen_books() |>
  group_by(book) |>
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text,
                                regex("^chapter [divxlc]",
                                      ignore_case = TRUE)))
  ) |>
  ungroup()

austen # As you can see in the below output, the austenr data has been now sorted by Austen's respective books and outlined by each chapter, delineated by the "linenumber" and "chapter" columns.
```

1.  **Tokenization**

    -   Use `unnest_tokens()` to tokenize text into individual words.

```{r}
# Here I filtering just for the 'Emma' book for the subsequent analysis
emma <- austen |>
  filter(book == "Emma")

# Using unnest_token & removing stopwords 
tidy_emma <- emma |>
  unnest_tokens(word, text) |>
  anti_join(stop_words)

tidy_emma # This output now shows the Emma book text without superfluous stopwords such as "a," "the," etc.
```

1.  **Sentiment Analysis**

    -   Filter joy/positive words from **each** of the three sentiment lexicons:

        -   `nrc` (joy)
        -   `afinn` (positive scores ≥ 1)
        -   `bing` (positive)

    -   Join each with *Emma*'s text and:

        -   Count word frequency.
        -   Filter for frequently occurring words (`n > 50`).
        -   Visualize using a **bar chart** (`ggplot2`) and a **word cloud** (`wordcloud`).

```{r}
# NRC joy sentiment analysis and visualization:
nrc_joy <- get_sentiments("nrc") |>
  filter(sentiment == "joy")

emma_nrc <- tidy_emma |>
  inner_join(nrc_joy, by = "word") |>
  count(word, sort = TRUE) |>
  filter(n > 50)

emma_nrc
```

```{r}
# Plot Visualization
ggplot(emma_nrc, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Joy Words in Emma (NRC)",
       x = "Word", y = "Frequency")

# You can see in the below plot that the most frequent "joy"-coded word in Emma is 'friend'
```

```{r}
# Joy Word Cloud Visualization
emma_nrc <- emma_nrc |>
  arrange(desc(n))

wordcloud(words = emma_nrc$word,
          freq = emma_nrc$n,
          max.words = 1000,
          colors = brewer.pal(8, "Dark2"))

# Same as the ggplot vis, we see 'friend' as the biggest word for the NRC joy wordcloud.
```

```{r}
# AFINN positive analysis and visualization
afinn <- get_sentiments("afinn")

emma_afinn <- tidy_emma |>
  inner_join(afinn, by = "word") |>
  filter(value >= 1) |>
  count(word, sort = TRUE) |>
  filter(n > 50)

emma_afinn
```

```{r}
# Plot visualization
ggplot(emma_afinn, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  labs(title = "Positive Words in Emma (AFINN)",
       x = "Word", y = "Frequency")

# we see 'dear' as the most frequent AFINN positive-coded word that appears in Emma.
```

```{r}
# Word Cloud visualization
emma_nrc <- emma_afinn |>
  arrange(desc(n))

wordcloud(emma_afinn$word, emma_afinn$n,
          max.words = 100,
          colors = brewer.pal(8, "Set2"))
# Same as shown via the ggplot vis, we see dear as the biggest word in the wordcloud for AFINN positive vocab.
```

```{r}
# Bing Positive Sentiment Analysis and Visualization
bing <- get_sentiments("bing")

emma_bing <- tidy_emma |>
  inner_join(bing, by = "word") |>
  filter(sentiment == "positive") |>
  count(word, sort = TRUE) |>
  filter(n > 50)

emma_bing
```

```{r}
# Plot Visualization
ggplot(emma_bing, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "purple") +
  coord_flip() +
  labs(title = "Positive Words in Emma (BING)",
       x = "Word", y = "Frequency")

# The most frequent Bing Positive-coded word in Emma is happy, closely followed by love.
```

```{r}
# Word Cloud Visualization
wordcloud(emma_bing$word, emma_bing$n,
          max.words = 1000,
          colors = brewer.pal(8, "Paired"))
# Similarly, we see here that happy is the most prominent Bing Positive-coded word in Emma.
```

1.  **Push to GitHub**

    -   Push your complete rendred qmd (`.html` file) to a **GitHub repository**.

    -   Your file should include:

        -   Data wrangling
        -   Sentiment filtering and joins
        -   Visualization code

    -   You must include **at least 5 meaningful commits** to document your workflow.

------------------------------------------------------------------------

## **Rubric 100 Points**

| Category                                       | Points  |
|------------------------------------------------|---------|
| Step 1: Data wrangling (linenumber, chapter)   | 10      |
| Step 2: Tokenization                           | 10      |
| Step 3a: NRC sentiment analysis + plots        | 10      |
| Step 3b: AFINN sentiment analysis + plots      | 10      |
| Step 3c: BING sentiment analysis + plots       | 10      |
| Code readability, structure, and comments      | 10      |
| Word cloud for each lexicon                    | 20      |
| GitHub submission                              |         |
| Commit history (at least 5 meaningful commits) | 20      |
| **Total**                                      | **100** |

Good Luck!
